---
title: "Analyses_by_Ohlert"
author: "Tim Ohlert"
date: "2/2/2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#load packages
```{r}
library(lmerTest)
library(tidyverse)
library(car)
library(visreg)
library(cAIC4)
library(MuMIn)
library(broom)
library(vegan)
library(ggthemes)
```


#load data
```{r}
df_annual <- read_csv("df_annual.csv")
```

```{r}
df_merge <- read_csv("df_merge.csv")
```

```{r}
df_site <- read_csv("df_site.csv")
```

```{r}
NutNetPheno_long <- read_csv("NutNetPheno_long.csv")
```


#RE: analysis. Totally simplified (amazing!!!) A three part analysis:
## STEP 1, show that there has been a directional, temporal shift towards "greening" and earlier springs since ~1980


```{r}
mod <- lmer(ChangeSOS ~ #change in green-up date from long-term avg.
              site * gs + #site * growing season (year)
              (1 | gs),
              data = df_merge)

Anova(mod)
summary(mod)


visreg(mod)
```




## STEP 2 go on a fishing expedition - does the *temporal* signature of phenology (e.g. the SLOPE of green-up date thru time) map onto any abiotic or biotic site-level characteristics?


```{r}
mod <- lmer(PhenoTrendXyr_sos_slope ~ #slope of green-up date change thru time
              abs(latitude) + #absolute value of latitude (since southern hemi sites!)
              Elevation + #elevation
              RainMAP + #long-term mean annual precip
              TempMAT + #long-term mean annual temp
              ndep_2016 + #modeled nitrogen deposition
              Richness + #alpha diversity
              FractionExotic +  #percent exotic species
              TempTrend36yr_2mo_sos_slope + #
              PrecipTrend36yr_2mo_sos_slope + #
              # ObsYrMeanEven + #NAs in this variable; so removed
              # SoilC + SoilN + SoilP + #NAs in this variable; so removed
              (1 | continent),
              data = df_merge)

Anova(mod)
summary(mod)

visreg(mod)
```


```{r}
a<-stepcAIC(mod,direction="backward", trace=TRUE, data=df_merge) #cant do w/singularities
```



## STEP 3, address interannual variability. Do phenological variations across years map onto annual weather patterns?

```{r}
mod <- lmer(ChangeSOS ~ #change in green-up date from site long-term average
              TempAnom_2mo_sos + #change in spring temp from site long-term average
              PrecipAnom_2mo_sos + #change in spring precip from site long-term average
              gs + #growing season (aka year)
              #as.factor(gs) + #year as a categorical factor
              (1 | continent),
            data = df_merge)
     #in this model you can use growing season as numeric but get singularity if you use as a factor


Anova(mod)
summary(mod)

visreg(mod)

```


```{r}
a<-stepcAIC(mod,direction="backward", trace=TRUE, data=df_merge) #cant do w/singularities
```











###################################################################
#permanova on step 2 model


## This is some code canibalized from Seabloom et al. 2019 Ecology paper.




# Compare adonis coefficients to linear model

```{r}
df <- df_merge[c("latitude", "Elevation", "RainMAP", "TempMAT", "ndep_2016", "Richness", "FractionExotic", "PhenoTrendXyr_sos_slope","PhenoTrendXyr_eos_slope","PhenoTrendXyr_gsl_slope", "PhenoTrendXyr_maxndvi_slope", "site", "continent", "TempTrend36yr_2mo_sos_slope", "PrecipTrend36yr_2mo_sos_slope")]
df <- df[complete.cases(df),]
df <- unique(df)


Y <- data.frame(df$PhenoTrendXyr_sos_slope, df$PhenoTrendXyr_eos_slope, df$PhenoTrendXyr_gsl_slope, df$PhenoTrendXyr_maxndvi_slope)#list of phenology response variables

nperm <-99
prm.test <- adonis(Y ~ abs(latitude) + #absolute value of latitude (since southern hemi sites!)
              Elevation + #elevation
              RainMAP + #long-term mean annual precip
              TempMAT + #long-term mean annual temp
              ndep_2016 + #modeled nitrogen deposition
              Richness + #alpha diversity
              FractionExotic #+  #percent exotic species
              # ObsYrMeanEven + #NAs in this variable; so removed
              # SoilC + SoilN + SoilP + #NAs in this variable; so removed
              #(1 | continent)
              , method="euclidean", na.rm = TRUE, data=df, permutations=nperm)
```



```{r}

perm <- how(nperm = 9)
setBlocks(perm) <- with(df, continent)
prm.test2 <- adonis2(Y ~ 
              TempTrend36yr_2mo_sos_slope + #
              PrecipTrend36yr_2mo_sos_slope + #
                abs(latitude) + #absolute value of latitude (since southern hemi sites!)
              Elevation + #elevation
              RainMAP + #long-term mean annual precip
              TempMAT + #long-term mean annual temp
              ndep_2016 + #modeled nitrogen deposition
              Richness + #alpha diversity
              FractionExotic #+  #percent exotic species

              # ObsYrMeanEven + #NAs in this variable; so removed
              # SoilC + SoilN + SoilP + #NAs in this variable; so removed
              #(1 | continent)
        , method="euclidean", data = df, permutations = perm)
```







```{r}
lm.test <- lm(Y[,1] ~ abs(latitude) + #absolute value of latitude (since southern hemi sites!)
              Elevation + #elevation
              RainMAP + #long-term mean annual precip
              TempMAT + #long-term mean annual temp
              ndep_2016 + #modeled nitrogen deposition
              Richness + #alpha diversity
              FractionExotic #+  #percent exotic species
              # ObsYrMeanEven + #NAs in this variable; so removed
              # SoilC + SoilN + SoilP + #NAs in this variable; so removed
              #(1 | continent)
              , data=df)
```




```{r}


x <- c("TempTrend36yr_2mo_sos_slope", "PrecipTrend36yr_2mo_sos_slope", "Latitude", "Elevation", "MAP", "MAT", "N.dep", "Richness", "FractionExotic",  "Residual", "Total")
z <- prm.test2[,3]
r2.data <- data.frame( x, z )
r2.data <- subset(r2.data, z<0.9)


ggplot(r2.data, aes(x,z))+
  geom_bar(stat="identity")+
  ylim(c(0,0.04))+
  ylab("Rate of phenology change (r2)")+
  xlab("")+
  theme_base()
  


```







##Are Richness and fraction exotic correlated?
```{r}
ggplot(df, aes(Richness, FractionExotic))+
  geom_point()+
         geom_smooth()+
         theme_base()

```










# This gets the slopes from the permanova

```{r}

cbind(coefficients(prm.test)[,1], coefficients(lm.test))
```

# This gets r2 values from a permanova
# Assemble results of permanova and plot r^2
# In order to plot the r^2 of each model term
```{r}

prm.tab.comp.j <- data.frame(prm.comp.j$aov.tab[,5:6])
```




